EldersAIR: Main analysis - Spirometry Outcomes

### 2021-05-06 main_analysis_final.Rmd
New RMDs for PM and Health paper results

---
title: 'EldersAIR: Main analysis - final version and manuscript results'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 26 April 2021"
output: pdf_document
---

@@

```{r}
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) 
```

@@

# Systolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

@@

# Diastolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_dia_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_dia) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# FVC
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fvc_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fvc_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# FEV1
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fev1_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fev1_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# PM2.5
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
pm_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = pm_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

pm_at_home_data <- analysis_data %>% 
  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

>>>>

---
title: 'EldersAIR: Main analysis final frameworks'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 5 May 2021"
output: pdf_document
---

@@

```{r}
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) %>% 
  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
  # Filter type isn't really feasible to use in the analysis due to this small sample size
  mutate(treatment_filter_type = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
                                         treatment_filter_type),
         treatment_filter_type = factor(treatment_filter_type2,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix")),
         filter_type = as.factor(filter_type),
         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
  mutate(treatment = as.character(treatment),
         perc_exp_quart = cut(kw_perc_expected,
                                  breaks = c(0, 25, 46, 71, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment == "Filter",
                                     perc_exp_quart, 
                                     treatment),
         treatment = factor(treatment,
                            levels = c("Placebo", "Education", "Filter")),
         filter_compliance = factor(filter_compliance,
                                    levels = c("Placebo", "Education",
                                               "0-25 compliance", "25-50 compliance",
                                               "50-75 compliance", "75-100 compliance"))) %>% 
  select(-treatment_filter_type2) 
```

@@

# Systolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))

itt_results_int <- lmer(outcome ~ treatment*gender +
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)

summary(int_emmeans)
```

@@

# Diastolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_dia_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_dia) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# FVC
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fvc_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fvc_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# FEV1
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fev1_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fev1_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# PM2.5
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
pm_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = pm_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, home) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

pm_at_home_data <- analysis_data %>% 
  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# Exposure-response analysis
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, pm_mean_at_home_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ pm_iqr + winter_id + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

tidy_results_combined <- rbind(tidy_results_primary, , 
                               )

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

### 2021-05-06 main_analysis_spiro.Rmd
New RMDs for PM and Health paper results

---
title: 'EldersAIR: Main analysis - Blood Pressure Outcomes'
author: "Ethan Walker"
date: "Started 6 May 2021; Updated 6 May 2021"
output: pdf_document
---

@@

analysis_data_2row <- data %>% 
  rename(outcome = outcome_var) %>%  
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# Systolic Blood Pressure ITT Results
```{r}
itt_health_function(analysis_data, "bp_sys", "Systolic Blood Pressure")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

@@

# Diastolic Blood Pressure ITT Results
```{r}
itt_health_function(analysis_data, "bp_dia", "Diastolic Blood Pressure")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

```{r}
itt_results_int <- lmer(outcome ~ treatment*gender +
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)

summary(int_emmeans)
```

>>>>

---
title: 'EldersAIR: Main analysis - Spirometry Outcomes'
author: "Ethan Walker"
date: "Started 6 May 2021; Updated 6 May 2021"
output: pdf_document
---

@@

analysis_data_2row <- data %>% 
  filter(spiro_status == "Acceptable") %>% 
  rename(outcome = outcome_var) %>%  
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

@@

# FVC ITT Results
```{r}
itt_health_function(analysis_data, "fvc", "FVC")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

@@

# FEV1 ITT Results
```{r}
itt_health_function(analysis_data, "fev1", "FEV1")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```
